---
title: "Control Structures in R"
author: "vishnusagar2021"
date: "1/28/2021"
output: html_document
---
# Control Structures in R - Loops in R
To control the execution of the expressions flow in R we use the control structures.
```{r}
#Control Structures in R - Loops in R
# There are eight types of control structures in R

# 8 loops

# *if()
# *if_else()
# *for()
# *nested loops()
# *while()
# *repeat and break()
# *next()
# *return()
```

## 1. if()
To carried out a task/command only `if` this condition is returned as *TRUE*
ie., condition is TRUE it will execute the command n when it FALSE it will not.
**Syntax**
'if (test_expression) {statement}'

```{r, echo=TRUE}
#eg.

x= sample(x= 1:100, size=1, replace=FALSE)     #data set  
x
#statement section........................if(test_expression){statement}
if(x%%2==0) {print(paste(x,"is an even number"))}                        
print
# paste, before statement print(paste,) used in the case of vector + character
```
```{r, echo=TRUE}
#eg2
values = 1:10
values
if(sample (values, 1) <5){print (paste (values,"is less than or equal to 5"))}
  

```

## if-else()
 True/False it can shows 2 outputs
 true for statement 1 or false for statement 2 one at a time
 
**syntax**
1. statement

2. if (test-expression) {statement} else {statemnt}

```{r, echo=TRUE}
# eg:
x= sample(x= 1:100, size=1, replace=FALSE)     #1. statement
x

if(x%%2 ==0) 
  {print(paste(x,"even number"))} else {print(paste(x, "odd number"))} #syntax

```
```{r, echo=TRUE}


```

